PANCHANGE_data_preprocessing

### 2020-08-17 PANCHANGE_data_preprocessing2020-07-15.Rmd
data export chunks

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(`Login ID`) %>% # Drop NAs
    distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category_full = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## RAMP
### Read in data
```{r ramp employment cols}
if(RAMP == TRUE){
ramp.employment.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
dim(ramp.employment.raw)
colnames(ramp.employment.raw)
}
```

@@

ramp.employment.raw.id <- ramp.employment.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.what_is_your_highest_level_of_education, #highest level of edcuation
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## Highest education (baseline, GLAD, EDGI, NBR, RAMP)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

@@

>>>>

ramp.dem.raw$externalDataReference <- ramp.dem.raw$`Login ID` #move Login ID to external data reference

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category_full = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## RAMP
### Read in data
```{r ramp employment cols}
if(RAMP == TRUE){
ramp.employment.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_coping_ramp.rds"))
dim(ramp.employment.raw)
colnames(ramp.employment.raw)
}
```

@@

ramp.employment.raw.id <- ramp.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## Highest education (baseline, GLAD, EDGI, NBR, RAMP)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

@@

```{r dat.coping summary}
summarytools::dfSummary(
  dat.coping,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

# Export data
```{r export data}
if(GLAD == TRUE){
saveRDS(object = dat.coping, file = paste0(data_path, "/glad.rds"))
}

if(EDGI == TRUE){
saveRDS(object = dat.coping, file = paste0(data_path, "/edgi.rds"))
}

if(NBR == TRUE){
saveRDS(object = dat.coping, file = paste0(data_path, "/nbr.rds"))
}

if(RAMP == TRUE){
saveRDS(object = dat.coping, file = paste0(data_path, "/ramp.rds"))
}
